More equivalences for Open pNets\thanks{This work was partially funded by the Associated Team FM4CPS between INRIA and ECNU, Shanghai}

### 2017-02-08 2016-Weak-FH-Bisim/WeakBisim.tex
\cite{Forte2016}
Proof of transitivity for FH bisim

\medskip
\ERIC{If we assume that everything is finite (states and transitions in the
open automata, and the predicates in $\mathcal{R}$, then it is easy to
prove that it is decidable whether a relation is a 
FH-bisimulation provided the logic of the predicates is decidable (see
\cite{Forte2016}). }

@@

The relation is built as follows: for each pair of state $s^1$, $s^3$, for each state 
$s^2$ such that $\mathcal{R}$ relates $s^1$ and $s^2$, and $\mathcal{R}'$ relates $s^2$ 
and $s^3$, we take the conjunction of the two predicates. The predicates for different 
values of $s^2$ are collected by a disjunction.
 we will show
that $\mathcal{R}''$ is an FH-bisimulation. Let
$(s^1,s^2|\Pred)\in\mathcal{R}$,  we have the followings since

\TODO{LUDO: I stopped rewriting here}
For any open transition $OT$ in $\mathcal{T}_1$:

@@

\inferrule*[myfraction=\reddottedrule]
{\{s^1_i\xrightarrow{a_i}_i s^{1'}_i\}^{i\in I_1}, % , s^1_i,s^{1'}_i\in
	%States(I_1)
	\{\xrightarrow{b_j}_j\}^{j\in J_1},\Pred_{OT},\Post_{OT}}
{s^1 \xrightarrow {v} s^{1'}}

there exist open transitions $OT_x^{x\in X} \subseteq \mathcal{T}_2$:

@@

%    \left( fresh \ \overline{a_i}, \overline{b_j}, v_x.\ \
\inferrule*[myfraction=\reddottedrule]
{\{s^{2}_i\xrightarrow{a_{i x}}_i s^{2}_{i x}\}^{i\in I_{2 x}},
	\{\xrightarrow{b_{j x}}_j\}^{j\in J_{2x}}, \Pred_{OT_x},\Post_{OT_x}}
{s^2 \xrightarrow {v_x} s^{2}_x}
%         \right)
\end{mathpar}

such that  $\forall x, J_1=J_{2x}, (s^{1'},s^{2}_x|\Pred_{target_x})\in 
\mathcal{R}$;
and  \\

$\Pred \land \Pred_{OT}\\
\hspace{1cm} \implies\!\!\! \bigvee_{x\in X}
\left( \forall j. b_j=b_{jx}  \Rightarrow \Pred_{OT_x}
\land v\!=\!v_x \land
\Pred_{target_x}\subst{\Post_{OT}}\subst{\Post_{OT_x}}\right)$

for any open transition $OT_x$, since
$(s^2,s^3|\Pred^\prime)\in\mathcal{R}$ there exist open transitions
$OT_{xy}^{x\in X, y\in Y} \subseteq \mathcal{T}_3$:

%    \left( fresh \ \overline{a_i}, \overline{b_j}, v_x.\ \
\inferrule*[myfraction=\reddottedrule]
{\{s^{3}_i\xrightarrow{a_{i x y}}_i s^{3}_{i xy}\}^{i\in I_{3 xy}},
	\{\xrightarrow{b_{j x y}}_j\}^{j\in J_{3xy}}, \Pred_{OT_{xy}},\Post_{OT_{xy}}}
{s^3 \xrightarrow {v_{xy}} s^{3}_{xy}}
%         \right)

\end{mathpar}
\TODO{LUDO: I have trouble doing computing the next line without intermediates: it is an 
iterative application of many FH bisim definitions with reordering of terms and 
manipulation of quantifiers}
such that  $\forall x, \forall y, J_1=J_{2x}=J_{3xy}, 
(s^{1'},s^{3}_{xy}|\Pred_{target_{xy}})\in \mathcal{R}$; and  \\
$\Pred\,' \land \Pred_{OT}\\
\hspace{1cm} \implies\!\!\! \bigvee_{x\in X}\bigvee_{y\in Y}
\left( \forall j. b_j=b_{jx}=b_{jxy}  \Rightarrow \Pred_{OT_{xy}}
\land v\!=\!v_{xy} \land
\Pred_{target_{xy}}\subst{\Post_{OT}}\subst{\Post_{OT_{xy}}}\right)$.

\TODO{the proof is incomplete because it does not deal with predicates and does not 
explicit what is the right $pred''$ $\implies$ TODO}

>>>>

\medskip
\ERIC{If we assume that everything is finite (states and transitions in the
open automata, and the predicates in $\mathcal{R}$, then it is easy to
prove that it is decidable whether a relation is a 
FH-bisimulation provided the logic of the predicates is decidable (see
\cite{henrio:Forte2016}). }

@@

The relation is built as follows: for each pair of state $s^1$, $s^3$, for each state 
$s^2$ such that $\mathcal{R}$ relates $s^1$ and $s^2$, and $\mathcal{R}'$ relates $s^2$ 
and $s^3$, we take the conjunction of the two predicates. The predicates for different 
values of $s^2$ are collected by a disjunction.

 We will show
that $\mathcal{R}''$ is an FH-bisimulation. Consider 
$(s^1,s^3|\Pred\,'')\in\mathcal{R}''$. Then there is a set of states of 
$\mathcal{T}_2$ relating $s^1$ and $s^3$, let $(s^2_p)^{p\in P}$ be this family.

For any $p\in P$ by definition of $\mathcal{R}''$,
$(s^1,s^2_p|\Pred_p)\in\mathcal{R}$,  and $(s^2_p,s^3|\Pred\,'_p)\in\mathcal{R}$. We have 
the 
following by definition of bisimulation.
For any open transition $OT$ in $\mathcal{T}_1$ originating from $s^1$:
Additionally we have $\Pred\,'' = \bigvee_{p\in P} \Pred_p\land\Pred_p\,'$.
\begin{mathpar}

\inferrule*[myfraction=\reddottedrule]
{\{s^1_i\xrightarrow{a_i}_i {s^{1}}'_i\}^{i\in I_1}, % , s^1_i,s^{1'}_i\in
	%States(I_1)
	\{\xrightarrow{b_j}_j\}^{j\in J_1},\Pred_{OT},\Post_{OT}}
{s^1 \xrightarrow {v} {s^{1}}'}

there exist open transitions $OT_{p x}^{x\in X} \subseteq \mathcal{T}_2$:

@@

%    \left( fresh \ \overline{a_i}, \overline{b_j}, v_x.\ \
\inferrule*[myfraction=\reddottedrule]
{\{s^{2}_{pi}\xrightarrow{a_{p i x}}_i s^{2}_{p i x}\}^{i\in I_{2 p x}},
	\{\xrightarrow{b_{j p x}}_j\}^{j\in J_{2px}}, \Pred_{OT_{px}},\Post_{OT_{px}}}
{s^2_p \xrightarrow {v_{px}} s^{2}_{px}}
%         \right)
\end{mathpar}

such that  $\forall x, J_1=J_{2px}, (s^{1'},s^{2}_{px}|\Pred_{target_{px}})\in 
\mathcal{R}$;
and  \\

$\Pred_p \land \Pred_{OT}\\
\hspace{1cm} \implies\!\!\! \bigvee_{x\in X}
\left( \forall j. b_j=b_{jpx}  \Rightarrow \Pred_{OT_{px}}
\land v\!=\!v_{px} \land
\Pred_{target_{px}}\subst{\Post_{OT}}\subst{\Post_{OT_{px}}}\right)$

for any open transition $OT_{px}$, since
$(s^2_p,s^3|\Pred\,^\prime_p)\in\mathcal{R}'$ there exist open transitions
$OT_{pxy}^{x\in X, y\in Y} \subseteq \mathcal{T}_3$:

@@

%    \left( fresh \ \overline{a_i}, \overline{b_j}, v_x.\ \
\inferrule*[myfraction=\reddottedrule]
{\{s^{3}_i\xrightarrow{a_{i p x y}}_i s^{3}_{i p x y}\}^{i\in I_{3 p x y}},
	\{\xrightarrow{b_{j p x y}}_j\}^{j\in J_{3pxy}}, \Pred_{OT_{pxy}},\Post_{OT_{pxy}}}
{s^3 \xrightarrow {v_{pxy}} s^{3}_{pxy}}
%         \right)

\end{mathpar}
such that  $\forall y, J_{2px}=J_{3pxy}, 
(s^{2}_x,s^{3}_{pxy}|\Pred_{target_{pxy}})\in \mathcal{R}'$; and  \\
$\Pred\,'_p \land \Pred_{OT_x}\\
\hspace{1cm} \implies\!\!\! \bigvee_{y\in Y}
\left( \forall j. b_{jpx}=b_{jpxy}  \Rightarrow \Pred_{OT_{pxy}}
\land v_{px}\!=\!v_{pxy} \land
\Pred_{target_{pxy}}\subst{\Post_{OT_{px}}}\subst{\Post_{OT_{pxy}}}\right)$.

This is verified for each $p\in P$. Overall,  we have a family of open transitions 
$OT_{pxy}^{p\in 
P, x\in X, 
y\in Y} \subseteq \mathcal{T}_3$ that should simulate \emph{OT}.

First, $\forall y, \forall x, \forall p, J_1=J_{2px}=J_{3pxy}, 
({s^{1}}',s^{3}_{pxy}|\Pred\,'_{target_{pxy}})\in \mathcal{R}''$. 
Indeed for any 
$p$, 
$x$, and 
$y$, $s^2_{px}$
relate ${s^{1}}'$ and $s^{3}_{pxy}$, indeed
$(s^{1'},s^{2}_{px}|\Pred_{target_{px}})\in \mathcal{R}$
 and $(s^{2}_x,s^{3}_{pxy}|\Pred_{target_{pxy}})\in \mathcal{R}'$. 
 More precisely, $s^{2}_{px} \in ({s^{2}_p}')^{p\in P'}$ where $({s^{2}_p}')^{p\in 
 P'}$ 
 is 
 the set of states relating ${s^{1}}'$ and $s^{3}_{pxy}$.
Consequently, for all $p$, $x$, $y$, $\Pred_{target_{px}}\land 
\Pred_{target_{pxy}}\implies 
\Pred\,'_{target_{pxy}}$ (this is one element of the  disjunction defining the 
predicate 
relating ${s^{1}}'$ and $s^{3}_{pxy}$ in the definition of $\mathcal{R}''$).

Second, we need \\
\begin{small}
$\Pred\,'' \land \Pred_{OT}\\
\hspace{1cm} \implies\!\!\! \bigvee_{x\in X}\bigvee_{y\in Y}\bigvee_{p\in P}
\left( \forall j. b_j=b_{jpxy}  \Rightarrow \Pred_{OT_{pxy}}
\land v\!=\!v_{pxy} \land
\Pred\,'_{target_{pxy}}\subst{\Post_{OT}}\subst{\Post_{OT_{pxy}}}\right)$.
\end{small}

We have, for all $p$
\begin{small}
$\Pred_p \land \Pred_{OT}\land\Pred'_p\\
\hspace{1cm} \implies\!\!\! \bigvee_{x\in X}
\left( \forall j. b_j=b_{jpx}  \Rightarrow \Pred_{OT_{px}}
\land v\!=\!v_{px} \land
\Pred_{target_{px}}\subst{\Post_{OT}}\subst{\Post_{OT_{px}}}\right)\land\Pred'_p\\
%
\hspace{1cm} \implies\!\!\! \bigvee_{x\in X}
\left( \forall j. b_j=b_{jpx}  \Rightarrow (\Pred_{OT_{px}}\land\Pred'_p)
\land v\!=\!v_{px} \land
\Pred_{target_{px}}\subst{\Post_{OT}}\subst{\Post_{OT_{px}}}\right)\\
%
\hspace{1cm} \implies\!\!\! \bigvee_{x\in X}
\big( \forall j. b_j=b_{jpx}  \Rightarrow (\bigvee_{y\in Y} 
\big( \forall j'. b_{j'px}=b_{j'pxy}  \Rightarrow \Pred_{OT_{pxy}}
\land v_{px}\!=\!v_{pxy}\\\hspace{3em}~ \land
\Pred_{target_{pxy}}\subst{\Post_{OT_{px}}}\subst{\Post_{OT_{pxy}}}\big))
\land v\!=\!v_{px} \land
\Pred_{target_{px}}\subst{\Post_{OT}}\subst{\Post_{OT_{px}}}\big)\\
%
\hspace{1cm} \implies\!\!\! \bigvee_{x\in X} \bigvee_{y\in Y}
\big( \forall j, j'. b_j=b_{jpx} \land b_{j'px}=b_{j'pxy}
\Rightarrow \big( 
\Pred_{OT_{pxy}}
\land v\!=v_{px}\!=\!v_{pxy}\\\hspace{3em}~ \land
\Pred_{target_{pxy}}\subst{\Post_{OT_{px}}}\subst{\Post_{OT_{pxy}}}
 \land
\Pred_{target_{px}}\subst{\Post_{OT}}\subst{\Post_{OT_{px}}}\big)\big)
$

\end{small}

Additionally, the Post renamings do not modify the predicates that only use non-renamed 
variables, and because of the domain of the renamings:\\
$\Pred_{target_{px}}\subst{\Post_{OT}}\subst{\Post_{OT_{px}}}\land 
\Pred_{target_{pxy}}\subst{\Post_{OT_{px}}}\subst{\Post_{OT_{pxy}}}\\
= 
\Pred_{target_{px}}\subst{\Post_{OT}}\subst{\Post_{OT_{px}}}\subst{\Post_{OT_{pxy}}}\land 
\Pred_{target_{pxy}}\subst{\Post_{OT}}\subst{\Post_{OT_{px}}}\subst{\Post_{OT_{pxy}}}\\
\implies\Pred\,'_{target_{pxy}}\subst{\Post_{OT}}\subst{\Post_{OT_{px}}}\subst{\Post_{OT_{pxy}}}\\
= \Pred\,'_{target_{pxy}}\subst{\Post_{OT}}\subst{\Post_{OT_{pxy}}} $

This allows us to conclude: using $\Pred\,'' = \bigvee_{p\in P} \Pred_p\land\Pred_p\,'$ 
and the assertions above we obtained the needed property.

\smallskip
Concerning the other direction of bisimulation, it is sufficient to notice that the role 
of $s^1$ and $s^3$ in the definition of $\mathcal{R}''$ is symmetrical, and thus the 
proof is similar.
